Relocating a cluster of earthquakes using a single station

### 2012-07-17 paper2/paper2.tex
some big changes done at GA

Earthquake location is important for many applications. It assists
our understanding of seismic activity by indicating regions of
relatively higher seismicity such as major lithospheric plate
boundaries \citep[e.g., ][]{dr_Sykes67a, dr_Isacks68a,
dr_Stein02a} or local active source zones
\citep[e.g., ][]{dr_Gutenberg44a, dr_Giardini03a}.
Earthquake locations are required for  magnitude determination
\citep[e.g., ][]{dr_Richter35a, dr_Gutenberg45a}, computing
moment tensors \citep[e.g., ][]{dr_Sipkin02a} and
seismological studies of the Earth's interior
\citep[e.g., ][]{dr_Spencer80a, dr_Kennett95a, dr_Curtis02a,
dr_Kennett04a}. They are needed to
understand strong motion and seismic attenuation
\citep[e.g., ][]{dr_Toro97a, dr_Campbell03a}
 and model earthquake hazard or risk
\citep[e.g., ][]{dr_Frankel00a, dr_Stirling02a, dr_Robinson06b}.
The accuracy required in earthquake location depends on the application.
For example, imaging the structure of a fracture system from microseismicity
requires greater detail than determining whether a $M_w=7.5$ earthquake
occurs offshore for tsunami warning purposes. This paper focuses
on reducing the uncertainty on the locations of a cluster of events, particularly
when they are recorded by a small number of stations.

Absolute location describes the location of an earthquake
with respect to some global reference such as latitude, longitude (or
easting/northing) and depth. Uncertainties associated with absolute locations
are influenced by source to station distances, the number of
stations and their geometry, quality of signal and accuracy of
the velocity model used in computing travel times. Errors in absolute location are typically of the
order of several km because they are susceptible to uncertainty in the velocity
structure along the entire path between the source and receiver.
For example, \citet{dr_Shearer99a} states that location uncertainty
in the ISC and PDE catalogues are generally around 25\,km horizontally and
at least 25\,km in depth.\footnote{Here the depth uncertainties of 25\,km assume the use
of depth dependent phases such as $pP$. Without such phases the uncertainty is higher.}
  These locations are created using global seismic networks.
\citet{dr_Bondar04a} demonstrate
that at the local scale, absolute locations are accurate to within 5\,km with a 95\%
confidence level when local networks meet a number of station
related criteria. Such errors are too large for many applications, particularly
those focussed on the study of micro seismicity and imaging
rupture surfaces from aftershock sequences.

Relative earthquake location involves locating a group of earthquakes with
respect to one another and was first introduced by \citet{dr_Douglas67a}
who developed the technique commonly known as joint hypocenter
determination.\footnote{\citet{dr_Douglas67a} originally used the term
joint epicentre determination. However, he was solving for hypocentre.}
In principle, relative locations can be
computed by differencing absolute locations. However, \citet{dr_Pavlis92a}
shows that inadequate knowledge of velocity structure leads to
systematic biases when relative positions are computed in this way.
This is confirmed by \citet{dr_Richards06a}
who state that inaccuracies in the travel time model are one of
two primary error sources in absolute earthquake location. The other
error source is erroneous arrival picks. To reduce errors from the travel
time model, relative location techniques compute locations
directly from travel time differences between two waveforms
\citep[e.g., ][]{dr_Ito85a, dr_Got94a, dr_Nadeau97a, dr_Waldhauser99a}.
By doing so, they remove errors
associated with unknown lateral velocity variations outside the
local region, because such variations influence all waveforms in
the same manner \citep{dr_Shearer99a}.

Reported uncertainties
for event locations from relative techniques are  around 15 to 75\,m in
local settings with good station coverage
\citep[e.g., ][]{dr_Ito85a, dr_Got94a, dr_Waldhauser99a,dr_Waldhauser08a}.
Here, `good coverage' implies multiple
stations distributed across a broad range of azimuthal directions.
Precise relative location techniques have been applied to a
diverse range of problems. For example, the location of active
fault planes \citep[e.g., ][]{dr_Deichmann92a, dr_Got94a,
dr_Waldhauser99a, dr_Waldhauser02a, dr_Shearer05a}, studying rupture mechanics
\citep[e.g., ][]{dr_Rubin99a, dr_Rubin02a}, interpreting magma
movement in volcanoes \citep[e.g., ][]{dr_Fremont87a} and monitoring
pumping induced seismicity \citep[e.g., ][]{dr_Lees98a,
dr_Ake05a}. \citet{dr_Poupinet84a},
\citet{dr_Bokelmann00a} and  \citet{dr_Rubin02b} apply relative
location techniques to identify earthquake doublets, events that have near identical
locations. \citet{dr_Poupinet84a} use these doublets to measure
temporal variation in crustal velocity, \citet{dr_Bokelmann00a}
use them to identify systematic temporal variations in seismic
anisotropy and \citet{dr_Rubin02b} describes how they are useful for
correcting time-dependent station delays.

@@

Coda refers to  later arriving waves in the seismogram that arise from
scattering  \citep{dr_Aki69a,dr_Snieder99a,dr_Snieder06a}.
Typically, the coda waves are ignored in most seismological
applications due to the complexity involved in modeling their
generation. In this
paper we develop an approach for locating earthquakes using coda waves.
\citet{dr_Snieder05a} demonstrate that the
coda of two earthquakes can be used to estimate the separation between
them. Their technique is based on the interference pattern between
the coda waves and is known as coda wave interferometry (CWI).
Unlike travel time techniques, CWI
 does not require multiple stations or good azimuthal coverage.
In fact, it is possible to obtain estimates of separation using a single
station \citep{dr_Robinson07b}. This makes CWI particularly interesting
for regions where station density is low such as intraplate areas.
Our technique can be used on coda waves alone or in combination with
travel times. We demonstrate that the use of coda waves improves
the location of a cluster of events when the recording situation
is poor and we show that it is possible to obtain accurate relative locations with as
little as one station.

\section{Theory}

\citet{dr_Snieder05a} introduce a CWI based estimator of source
separation $\delta_{CWI}$ between two earthquakes
\begin{equation}
\label{eq:Theory-dcwi_g_sigmat}
\delta_{CWI}^2 = g(\alpha,\beta)\sigma_\tau^2,
\end{equation}
where $\sigma_\tau$ is the standard deviation of the travel time perturbation
between the coda waves of two earthquakes,
and $\alpha$ and $\beta$ are the near source $P$ and $S$ wave velocities,
respectively. The function $g$ depends on the type of excitation
(explosion, point force, double couple) and on the direction of
source displacement relative to the point force or double
couple. For example, for two double couple sources displaced in the
fault plane,
\begin{equation}
\label{eq:LitReview-g-3d-dc}
g(\alpha,\beta) = 7
\frac{\left(\frac{2}{\alpha^6}+\frac{3}{\beta^6}\right)}{\left(\frac{6}{\alpha^8}+\frac{7}{\beta^8}\right)},
\end{equation}
whereas, for two point sources in a 2D acoustic medium
\begin{equation}
\label{eq:LitReview-g-2d-ps} g(\alpha,\beta) = 2 \alpha^2
\end{equation}
\citep{dr_Snieder05a}. The
$\sigma_\tau$ in eq. (\ref{eq:Theory-dcwi_g_sigmat}) is related to $R_{max}$, the maximum of the cross correlation between
the coda of the two waveforms and hence can be computed directly from the recorded data.
In this paper we use the autocorrelation
approach of \citet{dr_Robinson10b} when relating these parameters
and we apply a restricted time lag search when evaluating $R_{max}$. These extensions
to the original technique of \citet{dr_Snieder05a} increase the range of applicability
of CWI by 50\%.

\citet{dr_Robinson10b} show that coda waves provide
only probabilistic constraints on source separation and
introduce a Bayesian approach for describing the probability
of true separation given the CWI data. Their approach is
summarised by
\begin{equation}
\label{eq:-bayesian-noisyform}
P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN}) \propto P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)
\times P(\widetilde{\delta}_t)
\end{equation}
where  $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ is the posterior function
indicating the probability of true separation $\widetilde{\delta}_t$ given the noisy CWI separation
estimates $\widetilde{\delta}_{CWIN}$; $P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$
is the likelihood function representing the forward model between the
desired separation $\widetilde{\delta}_t$ and the data $\widetilde{\delta}_{CWIN}$; and
$P(\widetilde{\delta}_t)$ is the prior function accounting for all a-priori information.
The tilde above the separation parameters in eq. (\ref{eq:-bayesian-noisyform})
indicates the use of a wavelength normalised separation parameter
\begin{equation}
\label{eq-normalisation-eqn}
\widetilde{\delta} = \frac{\delta}{\lambda_{d}},
\end{equation}
which measures separation ($\delta$ = $\delta_{CWIN}$ or $\delta_t$) with
respect to dominant wavelength $\lambda_{d}$.
In this paper we consider a uniform prior which gives emphasis to the recorded data.
The algorithm for computing the noisy
likelihood $P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$ is derived by \citet{dr_Robinson10b}
and summarised in Appendix \ref{sec-Appendix-noisylikelihood}. With these two pieces
in place we can compute the posterior $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$
(or PDF) for the separation between any pair of events directly from their coda waves.

We seek a probability density function (PDF) which links
 individual pairwise posteriors $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$
and describes the location of multiple events with maximum identifying
 the most probable combination of locations.
More importantly however, the PDF will
quantify location uncertainty and provide information
on the degree to which individual events are constrained
 by the data. For convenience, we begin with three earthquakes having locations
$\mathbf{e}_1$, $\mathbf{e}_2$ and $\mathbf{e}_3$. Using a Bayesian formulation we write
\begin{equation}
\label{eq-Bayes-three-locations}
P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3| \mathbf{d}) \propto P( \mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)
\times P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3),
\end{equation}
where $P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \mathbf{d})$, $P(\mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$
and $P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ are the posterior,
likelihood and prior functions, respectively. In eq. (\ref{eq-Bayes-three-locations})
$\mathbf{d}$ represents observations that constrain the locations. They can be any combination
of travel times, geodetic information  or CWI separation estimations. For example, if coda waves are used
we have $P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \widetilde{\mitbf{\delta}}_{CWIN})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$
where $\widetilde{\mitbf{\delta}}_{CWIN}$ are the wavelength normalised separation estimates.
Alternatively, if we use CWI and travel time data we may write
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \widetilde{\mitbf{\delta}}_{CWIN}, \mathbf{t}_{DD})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN},\mathbf{t}_{DD}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$
where $\mathbf{t}_{DD}$ represent travel time differences.
In this derivation and in Sec. \ref{sec:benchmarking} we focus on the constraints imposed by coda waves, whereas in
Sec. \ref{sec:CalaverasLoc-CWIandTT} we demonstrate how
CWI and travel time data can be combined.

@@

We use synthetic examples in 2D with 50 earthquakes to test the performance of the optimisation
routine.

@@

Actual locations are created by
randomly selecting $\hat{x}$- and $\hat{y}$-coordinates such that $-50 \leq \hat{x} \leq 50$ and $-50 \leq \hat{y} \leq 50$.
These are indicated with triangles in Fig. \ref{fig-2D50eq-relocation_eg1}.
 We assume a local velocity of $v=3300\,$ms$^{-1}$ between all
event pairs and a dominant frequency of 2.5$\,$Hz to represent waveform data filtered between 1 and 5$\,$Hz.
The CWI data are defined by the positive bounded Gaussian with statistics
$\bar{\mu}_N$ and $\bar{\sigma}_N$. A hypothetical CWI mean is created by setting
\begin{equation}
\label{eq:hypothetical-CWI-mean-optichapt}
\bar{\mu}_N = \mu_1 \left( \widetilde{\delta}_t \right)
\end{equation}
using eq. (\ref{eq:mu1}). This assumption ensures that
the sample mean of hypothetical separation estimates is consistent with known CWI
biases \citep{dr_Robinson10b}. In example 1 we use $\bar{\sigma}_N=0.02$ between all event pairs.
Application of our optimisation procedure on the hypothetical CWI data yields the circles
in Fig. \ref{fig-2D50eq-relocation_eg1}. The optimisation does not lead to the correct
solution due to the addition of noise $\bar{\sigma}_N=0.02$ on the hypothetical
CWI data. The average coordinate error
is 2.0\,m which is small compared to the noise of $\bar{\sigma}_N=0.02$ which for
$v_s=3300$\,ms$^{-1}$ and $f_{dom}=2.5$\,Hz corresponds to roughly 25\,m.
\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_1.eps}
\caption{Example 1 - Synthetic relocation of 50 earthquakes in 2D using all constraints with noise $\bar{\sigma}_N=0.02$.
Actual and optimisation event locations
are shown in blue and red circles, respectively.}
\label{fig-2D50eq-relocation_eg1}
\end{figure}

\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_3.eps}
\caption{Example 3 - Synthetic relocation of 50 earthquakes in 2D using all constraints with noise
$\bar{\sigma}_N= 2 \sigma_{1to5Hz}(\delta_t)$.
 Actual and optimisation event locations
are shown in blue and red circles, respectively.}
\label{fig-2D50eq-relocation_eg3}
\end{figure}
\citet{dr_Robinson10b} demonstrates that the noise on CWI estimates is often larger than 0.02 and that it increases
with event separation. Consequently, example 1 is overly simplistic because we fix
$\bar{\sigma}_N=0.02$ for all pairs. In example 2 we increase the uncertainty and introduce a distance
dependance into the hypothetical $\bar{\sigma}_N$ by defining
$\bar{\sigma}_N=\sigma_{1to5Hz}(\delta_t)$, where $\sigma_{1to5Hz}(\delta_t)$ is the half-width of
the errorbars for a synthetic acoustic experiment
with filtering between 1 and 5\,Hz \citep[see Fig4(b) of ][]{dr_Robinson10b}.
Repeating the optimisation leads
to the circles in Fig. \ref{fig-2D50eq-relocation_eg3} which have an average coordinate error of 2.8\,m.

Conjugate gradient based optimisation techniques are susceptible to the presence of local minima.
This is because they use the slope of the target function to explore the solution space.
We explore the impact of local minima for our CWI location problem by beginning
the optimisation from 25 different randomly chosen starting positions. In this case we observe no differences
in the solution for neither examples 1 or 2.

We can draw three observations from the error structure in Figs. \ref{fig-2D50eq-relocation_eg1} and
\ref{fig-2D50eq-relocation_eg3}. Firstly, we note that the location errors depicted
by grey bars in each figure increase between examples 1 and 2 with the
introduction of larger noise. Secondly, the errors are larger for
events at greater distances from the center. This is because events near the center of the cluster are
constrained by links from all angles, whereas those
on the outside are moderated by links from a limited number of directions. This observation is analogous
to problems associated with poor azimuthal coverage in common triangulation problems such as individual earthquake location
from limited travel time data, or GPS positioning with few satellites. Our third observation is that the location errors appear to
be circular, despite our attempt to correct for rotational non-uniqueness with the local coordinate system.

The local coordinate system works by constraining the location of the first three
earthquakes. Earthquake 1 is fixed at the origin, earthquake 2 on the positive $\hat{x}$-axis and earthquake 3
has $\hat{y}>0$. As the number of events increases the strength of this constraint on later events
becomes weaker allowing small rotations of events with respect to each other.
That is, even though the rotational freedom of the cluster is in principal
removed by the constraints imposed on the events (see eqs. \ref{eq:coord-const1} to \ref {eq:coord-const4}) we observe that
in practice the effects of noise means that the rotational non-uniqueness
reappears. This is because the `easiest' way data noise can propagate is into
the direction which is least constrained by the data. It is the same phenomena as in
linear inversion where noise creates large spurious model changes in directions
of the eigenvectors with the smallest singular values \citep[e.g ][]{dr_Aster05a}.
Unfortunately, there is no obvious way to overcome this issue when using coda
waves alone since the constraints are based on separation. Fortunately however, combining coda waves with
measurements of travel times alleviates this problem and facilitate the removal of a local coordinate system
altogether.

Despite these observations about the error structure however, we gain confidence in the
optimisation procedure due to its stability for different starting locations and
because of the small average coordinate errors of 2.0\,m and 2.8\,m
for examples 1 and 2, respectively.

@@

A set of earthquakes and their corresponding coda wave estimates of separation can be thought of as
a network. In this network the earthquakes are nodes and the constraints are branches or links
which join them together. Synthetic examples 1 and 2 use 100\% direct linkage between event pairs. That is, there is a constraint
between each earthquake and all other events. In reality, we might expect that the separation between some
pairs will not be constrained by CWI data due to poor signal to noise ratio in the coda for common stations.
Obviously, the less stations that record an event the more likely it is that links
between it and other events will be broken.
In such cases the probabilistic distance constraint between a pair of
events may only exist indirectly through multiple pairs. In this section we
consider the impact of reduced linkage between event pairs. In example 3, we repeat example 2
using 90\%, 80\%, ..., 10\% of the links. As with the above examples, we undertake the
optimisation with 25 randomly chosen starting locations.

>>>>

Earthquake location is important for many applications. It assists
our understanding of seismic activity by indicating regions of
relatively higher seismicity such as lithospheric plate boundaries
\citep[e.g., ][]{dr_Sykes67a, dr_Isacks68a, dr_Stein02a} or local
active source zones \citep[e.g., ][]{dr_Gutenberg44a,
dr_Giardini03a}. Earthquake locations are required for  magnitude
determination \citep[e.g., ][]{dr_Richter35a, dr_Gutenberg45a},
computing moment tensors \citep[e.g., ][]{dr_Sipkin02a} and
seismological studies of the Earth's interior \citep[e.g.,
][]{dr_Spencer80a, dr_Kennett95a, dr_Curtis02a, dr_Kennett04a}. They
are needed to understand strong motion and seismic attenuation
\citep[e.g., ][]{dr_Toro97a, dr_Campbell03a}
 and model earthquake hazard or risk
\citep[e.g., ][]{dr_Frankel00a, dr_Stirling02a, dr_Robinson06b}.

The accuracy required in earthquake location depends on the
application. For example, imaging the structure of a fracture system
from microseismicity requires greater detail than determining
whether a $M_w=7.5$ earthquake occurs offshore for tsunami warning
purposes. This paper focuses on reducing the uncertainty on the
locations of a cluster of events when they are recorded by a small
number of stations.

Absolute location describes the location of an earthquake with
respect to a global reference such as latitude, longitude (or
easting/northing) and depth. Uncertainties associated with absolute
locations are influenced by source to station distances, the number
of stations and their geometry, quality of signal and accuracy of
the velocity model used in computing travel times. Uncertainties in
absolute location are typically of the order of several km because
they are susceptible to uncertainty in the velocity structure along
the entire path between the source and receiver. For example,
\citet{dr_Shearer99a} states that location uncertainty in the ISC
and PDE catalogues are generally around 25\,km horizontally and at
least 25\,km in depth.\footnote{Here the depth uncertainties of
25\,km assume the use of depth dependent phases such as $pP$.
Without such phases the uncertainty is higher.}
  These locations are created using global seismic networks.
\citet{dr_Bondar04a} demonstrate that at the local scale, absolute
locations are accurate to within 5\,km with a 95\% confidence level
when local networks meet a number of station related criteria. Such
errors are too large for many applications, particularly those
focussed on imaging rupture surfaces from aftershock sequences.

Relative earthquake location involves locating a group of
earthquakes with respect to one another and was first introduced by
\citet{dr_Douglas67a} who developed the technique commonly known as
joint hypocenter determination.\footnote{\citet{dr_Douglas67a}
originally used the term joint epicentre determination. However, he
was solving for hypocentre.} In principle, relative locations can be
computed by differencing absolute locations. However,
\citet{dr_Pavlis92a} shows that inadequate knowledge of velocity
structure leads to systematic biases when relative positions are
computed in this way. To reduce errors from unknown velocity
structure, relative location techniques compute locations directly
from travel time differences between two waveforms \citep[e.g.,
][]{dr_Ito85a, dr_Got94a, dr_Nadeau97a, dr_Waldhauser99a}. By doing
so, they remove errors associated with velocity variations outside
the local region, because such variations influence all waveforms in
the same manner \citep{dr_Shearer99a}.

Reported location uncertainties from relative techniques are  around
15 to 75\,m in local settings with good station coverage
\citep[e.g., ][]{dr_Ito85a, dr_Got94a,
dr_Waldhauser99a,dr_Waldhauser08a}. Here, `good coverage' implies
multiple stations distributed across a broad range of azimuthal
directions. Relative location techniques have been applied to a
diverse range of problems including: location of active fault planes
\citep[e.g., ][]{dr_Deichmann92a, dr_Got94a, dr_Waldhauser99a,
dr_Waldhauser02a, dr_Shearer05a}; studying rupture mechanics
\citep[e.g., ][]{dr_Rubin99a, dr_Rubin02a}; interpreting magma
movement in volcanoes \citep[e.g., ][]{dr_Fremont87a}; and
monitoring pumping induced seismicity \citep[e.g., ][]{dr_Lees98a,
dr_Ake05a}. \citet{dr_Poupinet84a}, \citet{dr_Bokelmann00a} and
\citet{dr_Rubin02b} apply relative location techniques to identify
earthquake doublets, events that have near identical locations.

@@

Coda refers to  later arriving waves in the seismogram that arise from
scattering  \citep{dr_Aki69a,dr_Snieder99a,dr_Snieder06a}.
Typically, the coda waves are ignored in most seismological
applications due to the complexity involved in modeling their
generation. In this
paper we develop an approach for locating earthquakes using coda waves.

\citet{dr_Snieder05a} demonstrate that the coda of two earthquakes
can be used to estimate the separation between them. Their
technique, known as coda wave interferometry (CWI), is based on the
interference pattern between the coda waves. Unlike travel time
techniques, CWI
 does not require multiple stations or good azimuthal coverage.
In fact, it is possible to obtain estimates of separation using a single
station \citep{dr_Robinson07b}. This makes CWI particularly interesting
for regions where station density is low such as intraplate areas.
Our technique can be used on coda waves alone or in combination with
travel times. We demonstrate that the use of coda waves improves
the location of a cluster of events when the recording situation
is poor and we show that it is possible to obtain accurate relative locations with as
little as one station.

\section{Theory}
\label{sec:theory}

\citet{dr_Snieder05a} introduce a CWI based estimator of source
separation $\delta_{CWI}$ between two earthquakes
\begin{equation}
\label{eq:Theory-dcwi_g_sigmat}
\delta_{CWI}^2 = g(\alpha,\beta)\sigma_\tau^2,
\end{equation}
where $\sigma_\tau$ is the standard deviation of the travel time perturbation
between the coda waves of two earthquakes,
and $\alpha$ and $\beta$ are the near source $P$ and $S$ wave velocities,
respectively. The function $g$ depends on the type of excitation
(explosion, point force, double couple) and on the direction of
source displacement relative to the point force or double
couple. For example, for two double couple sources displaced in the
fault plane,
\begin{equation}
\label{eq:LitReview-g-3d-dc}
g(\alpha,\beta) = 7
\frac{\left(\frac{2}{\alpha^6}+\frac{3}{\beta^6}\right)}{\left(\frac{6}{\alpha^8}+\frac{7}{\beta^8}\right)},
\end{equation}
whereas, for two point sources in a 2D acoustic medium
\begin{equation}
\label{eq:LitReview-g-2d-ps} g(\alpha,\beta) = 2 \alpha^2
\end{equation}
\citep{dr_Snieder05a}.

The $\sigma_\tau$ in eq. (\ref{eq:Theory-dcwi_g_sigmat}) is related
to $R_{max}$, the maximum of the cross correlation between the coda
of the two waveforms and hence can be computed directly from the
recorded data. In this paper we use the autocorrelation approach of
\citet{dr_Robinson11a} when relating these parameters and we apply a
restricted time lag search when evaluating $R_{max}$. These
extensions to the original technique of \citet{dr_Snieder05a}
increase the range of applicability of CWI by 50\%.

\citet{dr_Robinson11a} show that coda waves provide only
probabilistic constraints on source separation and introduce a
Bayesian approach for describing the probability of true separation
given the CWI data. Their approach is summarised by
\begin{equation}
\label{eq:-bayesian-noisyform}
P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN}) \propto P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)
\times P(\widetilde{\delta}_t)
\end{equation}
where  $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ is the posterior function
indicating the probability of true separation $\widetilde{\delta}_t$ given the noisy CWI separation
estimates $\widetilde{\delta}_{CWIN}$; $P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$
is the likelihood function representing the forward model between the
desired separation $\widetilde{\delta}_t$ and the data $\widetilde{\delta}_{CWIN}$; and
$P(\widetilde{\delta}_t)$ is the prior function accounting for all a-priori information.
The tilde above the separation parameters in eq. (\ref{eq:-bayesian-noisyform})
indicates the use of a wavelength normalised separation parameter
\begin{equation}
\label{eq-normalisation-eqn}
\widetilde{\delta} = \frac{\delta}{\lambda_{d}},
\end{equation}
which measures separation ($\delta$ = $\delta_{CWIN}$ or $\delta_t$)
with respect to dominant wavelength $\lambda_{d}$. In this paper we
consider a uniform prior which gives emphasis to the recorded data.
The algorithm for computing the noisy likelihood
$P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$ is derived by
\citet{dr_Robinson11a} and summarised in Appendix
\ref{sec-Appendix-noisylikelihood}. With these two pieces in place
we can compute the posterior
$P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ (or PDF) for the
separation between any pair of events directly from their coda
waves.

We seek a probability density function (PDF) which links
 individual pairwise posteriors $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$
to describe the location of multiple events with maximum identifying
 the most probable combination of locations.
More importantly however, the PDF will
quantify location uncertainty and provide information
on the degree to which individual events are constrained
 by the data.

For convenience, we begin with three earthquakes having locations
$\mathbf{e}_1$, $\mathbf{e}_2$ and $\mathbf{e}_3$. Using a Bayesian
formulation we write
\begin{equation}
\label{eq-Bayes-three-locations}
P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3| \mathbf{d}) \propto P( \mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)
\times P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3),
\end{equation}
where $P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \mathbf{d})$,
$P(\mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ and
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ are the posterior,
likelihood and prior functions, respectively. In eq.
(\ref{eq-Bayes-three-locations}) $\mathbf{d}$ represents
observations that constrain the locations. They can be any
combination of travel times, geodetic information  or CWI
separations. For example, if coda waves are used we have
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 |
\widetilde{\mitbf{\delta}}_{CWIN})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN}|\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3)$ where $\widetilde{\mitbf{\delta}}_{CWIN}$ are the
wavelength normalised separation estimates. Alternatively, if we use
CWI and travel time data we may write $P(\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3 | \widetilde{\mitbf{\delta}}_{CWIN},
\mathbf{\Delta}_{TT})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN},\mathbf{\Delta}_{TT}|\mathbf{e}_1,
\mathbf{e}_2, \mathbf{e}_3)$ where $\mathbf{\Delta}_{TT}$ represent
travel time differences. In this derivation and in Sec.
\ref{sec:benchmarking} we focus on the constraints imposed by coda
waves, whereas in Sec. \ref{sec:CalaverasLoc-CWIandTT} we
demonstrate how CWI and travel time data can be combined.

@@

We use synthetic examples in 2D and 3D with 50 earthquakes to test
the performance of the optimisation routine.

@@

Actual locations are created by
randomly selecting $\hat{x}$- and $\hat{y}$-coordinates such that $-50 \leq \hat{x} \leq 50$ and $-50 \leq \hat{y} \leq 50$.
These are indicated with triangles in Fig. \ref{fig-2D50eq-relocation_eg1}.
 We assume a local velocity of $v=3300\,$ms$^{-1}$ between all
event pairs and a dominant frequency of 2.5$\,$Hz to represent waveform data filtered between 1 and 5$\,$Hz.
The CWI data are defined by the positive bounded Gaussian with statistics
$\bar{\mu}_N$ and $\bar{\sigma}_N$. A hypothetical CWI mean is created by setting
\begin{equation}
\label{eq:hypothetical-CWI-mean-optichapt}
\bar{\mu}_N = \mu_1 \left( \widetilde{\delta}_t \right)
\end{equation}
using eq. (\ref{eq:mu1}). This assumption ensures that the sample
mean of hypothetical separation estimates is consistent with known
CWI biases \citep{dr_Robinson11a}. In example 1 we use
$\bar{\sigma}_N=0.02$ between all event pairs. Application of our
optimisation procedure on the hypothetical CWI data yields the
circles in Fig. \ref{fig-2D50eq-relocation_eg1}. The optimisation
does not lead to the correct solution due to the addition of noise
$\bar{\sigma}_N=0.02$ on the hypothetical CWI data. The average
coordinate error is 2.0\,m which is small compared to the noise of
$\bar{\sigma}_N=0.02$ which for $v_s=3300$\,ms$^{-1}$ and
$f_{dom}=2.5$\,Hz corresponds to roughly 25\,m.
\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_1.eps}
\caption{Example 1 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N=0.02$. Actual and
optimisation event locations are shown in triangles and circles,
respectively.} \label{fig-2D50eq-relocation_eg1}
\end{figure}

\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_3.eps}
\caption{Example 3 - Synthetic relocation of 50 earthquakes in 2D using all constraints with noise
$\bar{\sigma}_N= 2 \sigma_{1to5Hz}(\delta_t)$.
 Actual and optimisation event locations
are shown in triangles and circles, respectively.}
\label{fig-2D50eq-relocation_eg3}
\end{figure}
\citet{dr_Robinson11a} demonstrates that the noise on CWI estimates
is often larger than 0.02 and that it increases with event
separation. Consequently, example 1 is overly simplistic because we
fix $\bar{\sigma}_N=0.02$ for all pairs. In example 2 we increase
the uncertainty and introduce a distance dependance into the
hypothetical $\bar{\sigma}_N$ by defining
$\bar{\sigma}_N=\sigma_{1to5Hz}(\delta_t)$, where
$\sigma_{1to5Hz}(\delta_t)$ is the half-width of the errorbars for a
synthetic acoustic experiment with filtering between 1 and 5\,Hz
\citep[see Fig4(b) of ][]{dr_Robinson11a}. Repeating the
optimisation leads to the circles in Fig.
\ref{fig-2D50eq-relocation_eg3} which have an average coordinate
error of 2.8\,m.

Conjugate gradient based optimisation techniques are susceptible to
the presence of local minima. This is because they use the slope of
the target function to explore the solution space. We explore the
impact of local minima for our CWI location problem by beginning the
optimisation from 25 randomly chosen starting positions. We observe
no differences in the solution for either example.

We can draw three observations from the error structure in Figs.
\ref{fig-2D50eq-relocation_eg1} and \ref{fig-2D50eq-relocation_eg3}.
Firstly, the location errors depicted by grey bars increase between
examples 1 and 2 with the introduction of larger noise. Secondly,
the errors are larger for events at greater distances from the
center. This is because events near the center of the cluster are
constrained by links from all angles, whereas those on the outside
are moderated by links from a limited number of directions. This
observation is analogous to problems associated with poor azimuthal
coverage in common triangulation problems such as individual
earthquake location from limited travel time data, or GPS
positioning with few satellites. Our third observation is that the
location errors appear to be circular, despite our attempt to
correct for rotational non-uniqueness with the local coordinate
system.

The local coordinate system works by constraining the location of
the first three earthquakes. Earthquake 1 is fixed at the origin,
earthquake 2 on the positive $\hat{x}$-axis and earthquake 3 has
$\hat{y}>0$. As the number of events increases the strength of this
constraint on later events becomes weaker allowing small rotations
of events with respect to each other. That is, even though the
rotational freedom of the cluster is in principal removed by the
constraints imposed on the events (see eqs. \ref{eq:coord-const1} to
\ref {eq:coord-const4}) we observe that in practice the presence of
noise allows the rotational non-uniqueness to reappear. This is
because the `easiest' way data noise can propagate is into the
direction which is least constrained by the data. The same phenomena
is observed in linear inversion where noise creates large spurious
model changes in directions of the eigenvectors with the smallest
singular values \citep[e.g ][]{dr_Aster05a}. Unfortunately, there is
no obvious way to overcome this issue when using coda waves alone
since the constraints are based on separation. Fortunately however,
combining coda waves with measurements of travel times alleviates
this problem and facilitate the removal of a local coordinate system
altogether.

Despite our observations about the error structure however, we gain
confidence in the optimisation procedure due to its stability for
different starting locations and because of the small average
coordinate errors of 2.0\,m and 2.8\,m for examples 1 and 2,
respectively.

@@

A set of earthquakes and their corresponding coda wave estimates of
separation form a network. In this network the earthquakes are nodes
and the constraints are branches or links which join the nodes
together. Synthetic examples 1 and 2 use 100\% direct linkage
between event pairs. That is, there is a constraint between each
earthquake and all other events. In reality, we might expect that
the separation between some pairs will not be constrained by CWI
data due to poor signal to noise ratio in the coda for common
stations. Obviously, the less stations that record an event the more
likely it is that links between it and other events will be broken.
In such cases the probabilistic distance constraint between a pair
of events may only exist indirectly through multiple pairs. In this
section we consider the impact of reduced linkage between event
pairs. In example 3, we repeat example 2 using 90\%, 80\%, ..., 10\%
of the links. As with the above examples, we undertake the
optimisation with 25 randomly chosen starting locations.
